---
title: "Credit modelling"
author: "Victor Mandela"
date: "7/21/2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r include=FALSE}
loan_data <- readRDS("loan_data_ch1.rds")
```

You want to get an idea of the number, and percentage of defaults. Defaults are rare, so you always want to check what the proportion of defaults is in a loan dataset. 

The response variable loan_status, where 1 represents a default, and 0 represents non-default.

```{r}
# View the structure of loan_data
str(loan_data)

# Load the gmodels package 
library(gmodels)

# Call CrossTable() on loan_status
CrossTable(loan_data$loan_status)

# Call CrossTable() on grade and loan_status
CrossTable(x = loan_data$grade, y = loan_data$loan_status, prop.r = TRUE, prop.c = F, prop.t = F, prop.chisq = F)

```

## histogram for income

```{r}
# Create histogram of loan_amnt: hist_1
hist_1 <- hist(loan_data$loan_amnt)

# Print locations of the breaks in hist_1
hist_1$breaks

# Change number of breaks and add labels: hist_2
hist_2 <- hist(loan_data$loan_amnt, breaks = 200, xlab = "Loan amount", 
               main = "Histogram of the loan amount")
```

Bingo! Note that there are some high peaks at round values: 5000, 10000, 15000,… People tend to borrow round numbers. Makes sense, right?

## Outliers

There is a lot of blank space on the right-hand side of the plot. This is an indication of possible outliers. We will look at a scatterplot to verify this.

Note: If outliers are observed for several variables, it might be useful to look at bivariate plots. It's possible the outliers belong to the same observation. If so, there is even more reason to delete the observation because it is more likely that some information stored in it is wrong.

```{r}
# Plot the age variable
plot(loan_data$age, ylab = "Age")

# Save the outlier's index to index_highage
index_highage <- which(loan_data$age > 122)

# Create data set new_data with outlier deleted
new_data <- loan_data[-index_highage, ]

# Make bivariate scatterplot of age and annual income
plot(loan_data$age, loan_data$annual_inc, xlab = "Age", ylab = "Annual Income")

```

## Missing values

We can either:- delete, replace or keep missing values. We can keep missing values using coarse classification.

Coarse classification allows you to simplify your data and improve the interpretability of your model. Coarse classification requires you to bin your responses into groups that contain ranges of values. You can use this binning technique to place all NAs in their own bin.


```{r}
# Look at summary of loan_data
summary(loan_data$int_rate)

# Get indices of missing interest rates: na_index
na_index <- which(is.na(loan_data$int_rate))

# Remove observations with missing interest rates: loan_data_delrow_na
loan_data_delrow_na <- loan_data[-na_index, ]

# Make copy of loan_data
loan_data_delcol_na <- loan_data

# Delete interest rate column from loan_data_delcol_na
loan_data_delcol_na$int_rate <- NULL

```

```{r}
## Coarse classification
# Make the necessary replacements in the coarse classification example below 
loan_data$ir_cat <- rep(NA, length(loan_data$int_rate))

loan_data$ir_cat[which(loan_data$int_rate <= 8)] <- "0-8"
loan_data$ir_cat[which(loan_data$int_rate > 8 & loan_data$int_rate <= 11)] <- "8-11"
loan_data$ir_cat[which(loan_data$int_rate > 11 & loan_data$int_rate <= 13.5)] <- "11-13.5"
loan_data$ir_cat[which(loan_data$int_rate > 13.5)] <- "13.5+"
loan_data$ir_cat[which(is.na(loan_data$int_rate))] <- "Missing"

loan_data$ir_cat <- as.factor(loan_data$ir_cat)

# Look at your new variable using plot()
plot(loan_data$ir_cat)
```

## Modelling

To make your training and test sets, you should first set a seed using set.seed(). Seeds allow you to create a starting point for randomly generated numbers, so that each time your code is run the same answer is generated. The advantage of doing this in your sampling is that you or anyone else can recreate the exact same training and test sets by using the same seed.

```{r}
# Set seed of 567
set.seed(567)

# Store row numbers for training set: index_train
index_train <- sample(1:nrow(loan_data), 2 / 3 * nrow(loan_data))

# Create training set: training_set
training_set <- loan_data[index_train, ]

# Create test set: test_set
test_set <- loan_data[-index_train, ]
```

## creating confusion matrix

```{r}
# Create confusion matrix
conf_matrix <- table(test_set$loan_status, model_pred)

# Compute classification accuracy
(6092 + 349) / nrow(test_set)

# Compute sensitivity
349 / 1037
```

## logistic regression

```{r}
# Build a glm model with variable ir_cat as a predictor
log_model_cat <- glm(formula = loan_status ~ ir_cat, family = "binomial",
                     data = training_set)

# Print the parameter estimates 
log_model_cat

# Look at the different categories in ir_cat using table()
table(loan_data$ir_cat)
```

```{r}
# Build the logistic regression model
log_model_multi <- glm(loan_status ~ age + ir_cat + grade + loan_amnt +
                      annual_inc , family = "binomial", data = training_set)

# Obtain significance levels using summary()
summary(log_model_multi)
```

After having obtained all the predictions for the test set elements, it is useful to get an initial idea of how good the model is at discriminating by looking at the range of predicted probabilities. A small range means that predictions for the test set cases do not lie far apart, and therefore the model might not be very good at discriminating good from bad customers. With low default percentages, you will notice that in general, very low probabilities of default are predicted. It's time to have a look at a first model.

```{r}
# Build the logistic regression model
predictions_all_small <- predict(log_model_small, newdata = test_set, type = "response")

# Look at the range of the object "predictions_all_small"
range(predictions_all_small)
```

## big models expanding the range

```{r}
# Build the logistic regression model
log_model_full <- glm(loan_status ~ ., family = "binomial", data = training_set)

# Make PD-predictions for all test set elements using the the full logistic regression model
predictions_all_full <- predict(log_model_full, newdata = test_set, type = "response")

# Look at the predictions range
range(predictions_all_full)

```

## Specifying a cut-off

In the first argument, we are testing whether a certain value in the predictions-vector is bigger than 0.3. We get representation of "default" and "no default", respectively.

### comparing cut-offs

```{r}
# Fit the logit, probit and cloglog-link logistic regression models
log_model_logit <- glm(loan_status ~ age + emp_cat + ir_cat + loan_amnt,
                       family = binomial(link = logit), data = training_set)
log_model_probit <- glm(loan_status ~ age + emp_cat + ir_cat + loan_amnt,
                        family =  binomial(link = probit), data = training_set)
log_model_cloglog <- glm(loan_status ~ age + emp_cat + ir_cat + loan_amnt,
                         family = binomial(link = cloglog), data = training_set)

# Make predictions for all models using the test set
predictions_logit <- predict(log_model_logit, newdata = test_set, type = "response")
predictions_probit <- predict(log_model_probit, newdata = test_set, type = "response")
predictions_cloglog <- predict(log_model_cloglog, newdata = test_set, type = "response")

# Use a cut-off of 14% to make binary predictions-vectors
cutoff <- 0.14
class_pred_logit <- ifelse(predictions_logit > cutoff, 1, 0)
class_pred_probit <- ifelse(predictions_probit > cutoff, 1, 0)
class_pred_cloglog <- ifelse(predictions_cloglog  > cutoff, 1, 0)

# Make a confusion matrix for the three models
tab_class_logit <- table(true_val, class_pred_logit)
tab_class_probit <- table(true_val, class_pred_probit)
tab_class_cloglog <- table(true_val, class_pred_cloglog)

# Compute the classification accuracy for all three models
acc_logit <- sum(diag(tab_class_logit)) / nrow(test_set)
acc_probit <- sum(diag(tab_class_probit)) / nrow(test_set)
acc_cloglog <- sum(diag(tab_class_cloglog)) / nrow(test_set)
```

## Decision Trees

We begin with computing the gain of a tree using the GINI-measure.

```{r}

```

































